The \dealii Library, Version 9.5

### 2023-05-25 9.5/paper.tex
Fix grammar.

The current release includes numerous updates to the matrix-free
infrastructure, including:
\begin{itemize}
\item In release 9.3, we enabled parallel $hp$-operations in the matrix-free infrastructure.
The infrastructure did not work properly in the case that certain cells
did not get any degrees of freedom due to the usage of \texttt{FE\_Nothing}. This has been
fixed now. Furthermore, \texttt{FE\_Nothing} now also works together
with discontinuous Lagrange elements (i.e., with the \texttt{FE\_DGQ} class). Due to the popularity of \texttt{FE\_Nothing} as a mean to enable
or disable cells, we have introduced the new class
\texttt{ElementActivationAndDeactivationMatrixFree}, which wraps a \texttt{MatrixFree} object, only loops over all
active cells and optionally interprets faces between active and deactivated cells
as boundary faces. A use case is shown in~\cite{proell2023highly} in the context of powder-bed-fusion additive
manufacturing.
\item The matrix-free infrastructure allows to interleave cell loops with vector updates
by providing \texttt{pre}/\texttt{post} functions that are run on index ranges. This
feature is used, e.g., in \dealii to improve the performance of (preconditioned)
conjugate gradient solvers~\cite{kronbichler2022cg} as well as of relaxation and Chebyshev iterations (see Subsection~\ref{sec:lac}).
Up to release~9.3, the \texttt{pre}/\texttt{post} infrastructure was only supported for
continuous elements (cell loop); now, it also works for discontinuous
elements which also require face loops to assemble jump and penalty terms.
\item The operator \texttt{CellwiseInverseMassMatrix} now also efficiently
evaluates the inverse for coupling (dyadic) coefficients in the case of  multiple
components:
\begin{align*}
\left(v_i, D_{ij} u_j  \right)_{\Omega^{(K)}}
\quad  1\le i,j \le c,
\end{align*}
with $c$ being the number of components and $D\in \mathbb{R}^{c\times c}$ a tensorial
coefficient.
This is possible due to the tensor-product structure of the resulting element matrix:
\begin{align*}
M =  ( I_1 \otimes N^T) (D \otimes I_2 ) ( I_1 \otimes N),
\end{align*}
with $N$ being the tabulated shape functions and $I_1,I_2$ the appropriate identity matrices.
For square and invertible $N$, the inverse is explicitly given as:
\begin{align*}
M^{-1} =  ( I_1 \otimes N^{-1}) (D^{-1} \otimes I_2 ) ( I_1 \otimes N^{-T}).
\end{align*}
For hypercube-shaped cells, $N^{-1}$ has an explicit representation
again in terms of tensor products; for example, in 3d it can be
expressed as $N^{-1} = N_{1D}^{-1} \otimes N_{1D}^{-1} \otimes
N_{1D}^{-1}$, allowing the use sum factorization.
\end{itemize}

@@

\dealii has provided support for geometric multigrid methods%
\footnote{\dealii{} also supports algebraic multigrid methods
by interfacing to the external libraries \texttt{PETSc}
and \texttt{Trilinos}.}
for locally refined meshes nearly since its inception.
Traditionally, these were based on local-smoothing
methods (described many years later in \cite{JanssenKanschat2011}, see also
~\cite{ClevengerHeisterKanschatKronbichler2019}), and more
recently also global-coarsening algorithms~\cite{munch2022gc}. The global-coarsening
infrastructure, furthermore, allows to globally coarsen the polynomial degree ($p$-multigrid),
with specific applicability to $hp$-adaptive methods.
In the current release, we have added support for the case
that multigrid levels are given by non-nested meshes~\cite{adams2002evaluation, bittencourt2001nonnested, bramble1991analysis}. An example for such meshes is
presented in Figure~\ref{fig:nonnested}.

>>>>

The current release includes numerous updates to the matrix-free
infrastructure, including:
\begin{itemize}
\item In release 9.3, we enabled parallel $hp$-operations in the matrix-free infrastructure.
The infrastructure did not work properly in the case that certain cells
did not get any degrees of freedom due to the usage of \texttt{FE\_Nothing}. This has been
fixed now. Furthermore, \texttt{FE\_Nothing} now also works together
with discontinuous Lagrange elements (i.e., with the \texttt{FE\_DGQ} class). Due to the popularity of \texttt{FE\_Nothing} as a mean to enable
or disable cells, we have introduced the new class
\texttt{ElementActivationAndDeactivationMatrixFree}, which wraps a \texttt{MatrixFree} object, only loops over all
active cells and optionally interprets faces between active and deactivated cells
as boundary faces. A use case is shown in~\cite{proell2023highly} in the context of powder-bed-fusion additive
manufacturing.
\item The matrix-free infrastructure allows interleaving cell loops with vector updates
by providing \texttt{pre}/\texttt{post} functions that are run on index ranges. This
feature is used, e.g., in \dealii to improve the performance of (preconditioned)
conjugate gradient solvers~\cite{kronbichler2022cg} as well as of relaxation and Chebyshev iterations (see Subsection~\ref{sec:lac}).
Up to release~9.3, the \texttt{pre}/\texttt{post} infrastructure was only supported for
continuous elements (cell loop); now, it also works for discontinuous
elements which also require face loops to assemble jump and penalty terms.
\item The operator \texttt{CellwiseInverseMassMatrix} now also efficiently
evaluates the inverse for coupling (dyadic) coefficients in the case of  multiple
components:
\begin{align*}
\left(v_i, D_{ij} u_j  \right)_{\Omega^{(K)}}
\quad  1\le i,j \le c,
\end{align*}
with $c$ being the number of components and $D\in \mathbb{R}^{c\times c}$ a tensorial
coefficient.
This is possible due to the tensor-product structure of the resulting element matrix:
\begin{align*}
M =  ( I_1 \otimes N^T) (D \otimes I_2 ) ( I_1 \otimes N),
\end{align*}
with $N$ being the tabulated shape functions and $I_1,I_2$ the appropriate identity matrices.
For square and invertible $N$, the inverse is explicitly given as:
\begin{align*}
M^{-1} =  ( I_1 \otimes N^{-1}) (D^{-1} \otimes I_2 ) ( I_1 \otimes N^{-T}).
\end{align*}
For hypercube-shaped cells, $N^{-1}$ has an explicit representation
again in terms of tensor products; for example, in 3d it can be
expressed as $N^{-1} = N_{1D}^{-1} \otimes N_{1D}^{-1} \otimes
N_{1D}^{-1}$, allowing the use sum factorization.
\end{itemize}

@@

\dealii has provided support for geometric multigrid methods%
\footnote{\dealii{} also supports algebraic multigrid methods
by interfacing to the external libraries \texttt{PETSc}
and \texttt{Trilinos}.}
for locally refined meshes nearly since its inception.
Traditionally, these were based on local-smoothing
methods (described many years later in \cite{JanssenKanschat2011}, see also
~\cite{ClevengerHeisterKanschatKronbichler2019}), and more
recently also global-coarsening algorithms~\cite{munch2022gc}. The global-coarsening
infrastructure, furthermore, allows globally coarsening the polynomial degree ($p$-multigrid),
with specific applicability to $hp$-adaptive methods.
In the current release, we have added support for the case
that multigrid levels are given by non-nested meshes~\cite{adams2002evaluation, bittencourt2001nonnested, bramble1991analysis}. An example for such meshes is
presented in Figure~\ref{fig:nonnested}.
